Getting Started with the Graded Response Model (GRM): A gentle introduction and tutorial in R

### 2024-01-17 draft_paper.qmd
All steps done!

---
title: "Getting Started with the Graded Response Model (GRM): A gentle introduction and tutorial in R"
shorttitle: "A Graded Response Model Tutorial"
author:
  - name: Rizqy Amelia Zein
    corresponding: true
    orcid: 0000-0001-7840-0299
    email: amelia.zein@psikologi.unair.ac.id
    affiliations:
      - name: Universitas Airlangga
        department: Department of Psychology
        address: Jalan Airlangga 4-6
        city: Surabaya
        region: Jawa Timur, Indonesia
        postal-code: "60286"
      - name: Ludwig-Maximilians-Universität
        department: Department of Psychology
        address: Leopoldstraße 13
        city: Munich
        postal-code: "80802"
  - name: Hanif Akhtar
    orcid: 0000-0002-1388-7347
    affiliations:
      - name: ELTE Eötvös Loránd University
        department: Doctoral School of Psychology
        country: Hungary
      - name: Universitas Muhammadiyah Malang
        department: Faculty of Psychology
        city: Malang
        region: Jawa Timur, Indonesia
abstract: "This tutorial paper introduces the Graded Response Model (GRM), a tool for testing measurement precision under the Item Response Theory (IRT) paradigm. Addressing common problems of measurement imprecision and lack of construct validity, the tutorial guides researchers through a one-dimensional GRM analysis in the R environment, using psych, mirt, and ggmirt packages. GRM is specifically designed to examine the psychometric properties of psychological scales with polytomous items. The tutorial illustrates the procedure using data from the Open Psychometrics Database on the Right-Wing Authoritarianism (RWA) scale, outlining the theoretical underpinnings of GRM, and steps for data preparation, model fitting, interpretation of results, and dealing with common issues and anomalies that may typically arise in the process."
keywords:
  - graded response model
  - item response theory
  - mirt
  - ggmirt
  - psych
  - R
author-note:
  disclosures:
    conflict-of-interest: "We declare we have no competing interests."
    financial-support: "The first author receives a PhD scholarship from the German Academic Exchange Service (Deutscher Akademischer Austauschdienst, DAAD)."
lang: en
bibliography: references.bib
format:
  apaquarto-docx: default
editor: visual
execute: 
  warning: false
  error: false
  output: false
  echo: true
---

@@

The output of this function is provided in {apatb-table5}, which shows fit statistics for an individual item. All items have very good (low) *RMSEA* values but normally, we want the *p* values of the scaled $\chi$^2^ to be nonsignificant, implying there is no discrepancy between the observed response pattern with what the model predicts. In {apatb-table5}, only three items here are nonsignificant (Q2, Q4, and Q12), which strengthens our suspicion that the RWA items might not be locally independent.

@@

The output of this code is {apafg-fig6}, and according to this, the RWA scale as a whole is informative for measuring individuals with a wide range of RWA levels, i.e., between -2*SD* to +4*SD*.

@@

```{r}
m_rel <- marginal_rxx(fit)
```

lhalalala

@@

{apafg-fig7}

```{r}
theta_se <- fscores(fit, full.scores.SE = T) # Extracting the estimated theta score of each participant.
e_rel <- empirical_rxx(theta_se) # Then use the estimated theta to calculate empirical reliability.
```

lhalalala

```{r}
omega(rwa)
```

>>>>

---
title: "Getting Started with the Graded Response Model (GRM): A gentle introduction and tutorial in R"
shorttitle: "A Graded Response Model Tutorial"
author:
  - name: Rizqy Amelia Zein
    corresponding: true
    orcid: 0000-0001-7840-0299
    email: amelia.zein@psikologi.unair.ac.id
    affiliations:
      - name: Universitas Airlangga
        department: Department of Psychology
        address: Jalan Airlangga 4-6
        city: Surabaya
        region: Jawa Timur, Indonesia
        postal-code: "60286"
      - name: Ludwig-Maximilians-Universität
        department: Department of Psychology
        address: Leopoldstraße 13
        city: Munich
        postal-code: "80802"
  - name: Hanif Akhtar
    orcid: 0000-0002-1388-7347
    affiliations:
      - name: ELTE Eötvös Loránd University
        department: Doctoral School of Psychology
        country: Hungary
      - name: Universitas Muhammadiyah Malang
        department: Faculty of Psychology
        city: Malang
        region: Jawa Timur, Indonesia
abstract: "This tutorial paper introduces the Graded Response Model (GRM), a tool for testing measurement precision under the Item Response Theory (IRT) paradigm. Addressing common problems of measurement imprecision and lack of construct validity, the tutorial guides researchers through a unidimensional GRM analysis in the R environment, using psych, mirt, and ggmirt packages. GRM is specifically designed to examine the psychometric properties of psychological scales with polytomous (Likert-style) items. The tutorial illustrates the procedure using data from the Open Psychometrics Database on the Right-Wing Authoritarianism (RWA) scale, outlining the theoretical underpinnings of GRM, and steps for data preparation, model fitting, interpretation of results, and dealing with common issues and anomalies that may typically arise in the process."
keywords:
  - graded response model
  - item response theory
  - mirt
  - ggmirt
  - psych
  - R
author-note:
  disclosures:
    conflict-of-interest: "We declare we have no competing interests."
    financial-support: "The first author receives a PhD scholarship from the German Academic Exchange Service (Deutscher Akademischer Austauschdienst, DAAD)."
lang: en
bibliography: references.bib
format:
  apaquarto-docx: default
editor: visual
execute: 
  warning: false
  error: false
  output: false
  echo: true
---

@@

The output of this function is provided in {apatb-table5}, which shows fit statistics for an individual item. All items have very good (low) *RMSEA* values but normally, we want the *p* values of the scaled $\chi$^2^ to be nonsignificant, implying that there is no discrepancy between the observed response pattern with what the model predicts. In {apatb-table5}, only three items here are nonsignificant (Q2, Q4, and Q12), which strengthens our suspicion that the RWA items might not be locally independent.

@@

The output of this code is {apafg-fig6}, and according to this, the RWA scale as a whole is informative for measuring a group of individuals with a wide range of RWA levels, i.e., between -2*SD* to +4*SD*. However, the RWA scale is not optimal for measuring individuals with RWA levels below or beyond this range.

@@

The heart of an IRT analysis is to estimate the precision of the measure, which can vary given different levels of   $\theta$. *mirt* provides two strategies to compute reliability that differs in their assumption regarding the distribution of $\theta$. First, readers may calculate marginal reliability by running this simple command:

```{r}
marginal_rxx(fit)
```
```{r}
#| include: false

m_rel <- marginal_rxx(fit)
m_rel = round(m_rel, digits=3)
```

The output of the above code is `r m_rel`, which indicates that the RWA scale as a whole is reliable, assuming that the underlying $\theta$ distribution follows the Gaussian or normal distribution [@chalmersAnswerDifferenceEmpirical2019]. Since reliability can vary across different levels of $\theta$, readers can also visualize the scale's reliability given the levels of $\theta$ by running the function of *ggmirt* package, as follows:

@@

The output of this code is {apafg-fig7}, which shows that the RWA scale can measure individuals with $\theta$ levels between -2*SD* and +4*SD* with sufficient reliability, i.e., *r*~xx~ $\geq$ 0.75.

Second, *mirt* can also calculate the RWA scale overall reliability using $\theta$ predicted by the model. To compute empirical reliability, one needs to calculate estimated $\theta$ value of each participant first, and then, calculate reliability based on these values. To do this, readers can run these two lines of codes:

```{r}
theta_se <- fscores(fit, full.scores.SE = TRUE) # Extracting the estimated theta score of each participant.
empirical_rxx(theta_se) # Then use the estimated theta to calculate empirical reliability.
```
```{r}
#| include: false

e_rel <- empirical_rxx(theta_se)
e_rel = round(e_rel, digits=3)
```

The `fscores()` function is used to compute the estimated $\theta$ value for each participant predicted by the `fit` model. The `full.scores.SE` is set to `TRUE` because the package uses the $\theta$ value and standard error of each participant (*SE*~$\theta$~) to compute the empirical reliability. We store the estimated $\theta$ and its *SE*~$\theta$~ in a new data frame, namely `theta_se`, and then run `empirical_rxx()` function to calculate empirical reliability. The output of these codes is `r e_rel`, which is very close to marginal reliability. However, it is very important to note that while marginal and empirical reliability aim to quantify the scale's overall reliability, they base the calculations on different assumptions. Marginal reliability assumes that $\theta$ is normally distributed, and empirical reliability uses $\theta$ distribution predicted by the model, which still contains measurement errors but more realistic.

It is very important to note that our model has a poor fit, as we showed in step 4. In practice, we recommend that readers address the model fitting issues before interpreting reliability or drawing substantive conclusions based on these estimates.

To paint a more comprehensive picture, we can compare IRT-based reliability calculations with unweighted sum-score based reliability, such as Cronbach's $\alpha$ and McDonald's $\omega$. To do this, we can use `omega()` function provided by *psych* package [@revellePsychProceduresPsychological2023], which can be done by running this command:

```{r}
omega(rwa)
```
```{r}
#| include: false

om <- omega(rwa)
alpha <- om$alpha
alpha = round(alpha, digits = 3)
oh <- om$omega_h
oh = round(oh, digits = 3)
```

The output of `omega()` function includes many information but we only focus on Cronbach's $\alpha$ and McDonald's $\omega$~h~. Both Cronbach's $\alpha$ ($\alpha$ = `r alpha`) and McDonald's $\omega$ ($\omega$~h~ = `r oh`) shows that the RWA scale is overall internally consistent.
